When, why, and how to avoid, prefer, or tolerate inbreeding

### 2018-04-01 presentation/inbreeding.Rmd
More slides

## When to inbreed versus avoid inbreeding?

```{r, echo = FALSE, fig.width = 3}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}
inputPanel(
  sliderInput("rr", label = "Relatedness", width="100%",
              min = 0.0, max = 1, value = 0, step = 0.05), 
  sliderInput("dd", label = "Inbreeding Depression", width="100%",
              min = 0.0, max = 1, value = 0, step = 0.1)
)
renderPlot({
    par(bg = 'transparent', mar = c(0.5, 0.5, 0.5, 0.5));
    plot(x = -1:1, y = -1:1, xaxt = "n", yaxt = "n", xlab = "", ylab = "Fitness", 
         type = "n", xlim = c(-1, 1), ylim = c(-1, 1), bty = "n");
    val <- seq(from = (-1 * pi)/4, to = (1 * pi)/4, length = 1000);
    hx  <- cos(val);
    hy  <- sin(val);
    fva <- seq(from = 0, to = 1, length = 1000);
    fit <- 0.5*(1 + input$rr)*(1 - input$dd);
    sqd <- (fva - fit)*(fva - fit);
    clf <- which(sqd == min(sqd));
    fx  <- cos(val[clf]);
    fy  <- sin(val[clf]);
    points(x = hx, y = hy, lwd = 2, type = "l");
    points(x = seq(from = 0, to = 1, length = 100), y = rep(0, 100), 
           type = "l", lty = "dotted", col = "red");
    ia_box <- mbox(x0 = -1.1, x1 = -0.1, y0 = -0.9, y1 = -1.1);
    polygon(x=ia_box$x, y=ia_box$y, lwd=3, border="black", col="black");
    ip_box <- mbox(x0 = 0.1, x1 = 1.1, y0 = -0.9, y1 = -1.1);
    polygon(x=ip_box$x, y=ip_box$y, lwd=3, border="black", col="black");
    points(x = seq(from = -1, to = 1, length = 100), y = rep(-0.85, 100), 
           lwd = 6, type = "l");
    points(x = rep(0, 100), y = seq(from = -0.85, to = 0, length = 100), 
           lwd = 6, type = "l");
    text(x = -0.6, y = -1, labels = "Avoid inbreeding", cex = 1.75, col = "red");
    text(x = 0.6, y = -1, labels = "Inbreeding", cex = 1.75, col = "white");
    text(x = 0.8, y = 0.75, labels = "n", cex = 3);
    text(x = 0.8, y = -0.75, labels = "0", cex = 3);
    text(x = -1, y = 0, labels = expression(frac(n,2)), cex = 3);
    arrows(x0 = 0, x1 = fx, y0 = 0, y1 = fy, lwd = 3, col = "black");
    arrows(x0 = 0, x1 = -0.9, y0 = 0, y1 = 0, lwd = 3, col = "red");
    points(x = 0, y = 0, cex = 5, pch = 20, col = "black");
    points(x = 0, y = 0, cex = 3, pch = 20, col = "red");
})
```

Is inclusive fitness from avoiding inbreeding (left side) less than inbreeding (right side)?

$$\frac{n}{2} < \frac{n}{2} (1 + r)(1 - \delta)$$

More generally, is inclusive fitness from inbreeding (right) greater than inclusive fitness from outbreeding (left)?

$$\delta < \frac{r}{1+r}$$

@@

@@

@@

<br><br><br><br><br><br>

@@

@@

@@

@@

## Other stuff to do

Make point that it's hard to detect through a change in kinship (and subject to bias)

Make point that pre, post and polyandry evolve given strong ID and no costs

Make point that opportunity for post-cop greatly increases polyandry selection

Make point that selection tends to go with the least costly IA route

>>>>

<center>
![](images/fem_choice_7.png)

@@

**Female threshold:**

$$1 > (1+r)(1-\delta)$$

**Male threshold:**

$$r > (1+r)(1-\delta)$$

@@

![](images/hist_blank.png)
<br><br><br><br><br><br><br><br>

@@

<br><br><br><br><br>

@@

<br><br><br><br><br><br><br><br><br><br><br>

@@

<center>
![](images/pre_post_inds.png)

## Effect of post-copulatory selection on polyandry {.smaller}

@@

<span style="font-size: 18px;">
Duthie, AB et al. (2018) *J. Evol. Biol.* [**31**:31-45](https://bradduthie.github.io/DuthieEtAl2017.pdf).
</span>

## Effect of post-copulatory selection on polyandry {.smaller}

@@

<span style="font-size: 18px;">
Duthie, AB et al. (2018) *J. Evol. Biol.* [**31**:31-45](https://bradduthie.github.io/DuthieEtAl2017.pdf).
</span>

## Effect of post-copulatory selection on polyandry {.smaller}

<center>
![](images/pre_post_cost_comp.png)
</center>

<span style="font-size: 18px;">
Duthie, AB et al. (2018) *J. Evol. Biol.* [**31**:31-45](https://bradduthie.github.io/DuthieEtAl2017.pdf).
</span>

## Post-copulatory inbreeding avoidance conclusions

1. Post-copulatory inbreeding avoidance can [greatly strengthen for polyandry](https://bradduthie.github.io/DuthieEtAl2017.pdf#page=8)
2. How inbreeding is avoided is affected by the [relative cost of pre-copulatory versus post-copulatory choice](https://bradduthie.github.io/DuthieEtAl2017.pdf#page=10)

<br><br><br><br><br><br><br><br><br>

<span style="font-size: 18px;">
Duthie, AB et al. (2018) *J. Evol. Biol.* [**31**:31-45](https://bradduthie.github.io/DuthieEtAl2017.pdf).
</span>

## Individual-based models: general conclusions

## Individual-based models: general conclusions

<hr>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf). <br>

</span>

## Individual-based models: general conclusions

<hr>
<span style="font-size: 18px;">
2. Duthie, AB et al. (2016) *Evolution* [**70**:1927-1943](https://bradduthie.github.io/DuthieEtAl2016b.pdf). <br>
</span>

## Individual-based models: general conclusions

<hr>
<span style="font-size: 18px;">
3. Duthie, AB et al. (2018) *J. Evol. Biol.* [**31**:31-45](https://bradduthie.github.io/DuthieEtAl2017.pdf). <br>
</span>

## Modulating parental investment and inbreeding
